Fast and Error-Adaptive Influence Maximization based on Count-Distinct Sketches

### 2021-04-02 paper.tex
Update on Overleaf.

\IEEEtitleabstractindextext{%
\begin{abstract}
Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation methods are frequently used for IM. In addition to these high-quality, yet expensive approximation algorithms, lightweight, sketch-based approaches, which do not simulate the process exactly, have been proposed in the literature to cope with the scale of today's networks. %As expected, the quality of the seed sets for the latter is inferior to that of the former. 
In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and the recently proposed hash-based fused sampling. To estimate the number of influenced vertices throughout a diffusion, we use per-vertex Flajolet-Martin sketches where each sketch corresponds to a sampled subgraph. To efficiently simulate the diffusions, the reach-set cardinalities of a single vertex are stored in memory in a consecutive fashion. This allows the proposed algorithm to find the number of influenced vertices in a single simulation step of different simulations at once. In addition, thanks to their efficiency, and the scalability of our parallel implementation, we can rebuild the sketches whenever it is necessary. For a faster IM kernel, we rebuild the sketches only after observing estimation errors above a given threshold. Our experimental results show that the proposed method yields high-quality seed sets while being up-to $1.7\times$--$27.9\times$ faster on average than a state-of-the-art approximation algorithm. In addition, it is $3.5\times$--$11.1\times$ faster than a sketch-based approach while producing seed sets with $3\%$--$12\%$ better influence scores.
\end{abstract}
}

@@

The study of efficient information/influence propagation in networks becomes an important area of research with several applications in many fields, such as viral marketing~\cite{leskovec2007dynamics, trusov2009effects}, social media analysis~\cite{zeng2010social, moreno2004dynamics}, and recommendation systems~\cite{lu2012recommender}.
As the study of these networks is imperative for educational, political, economic, and social purposes, a high-quality seed set to initiate the diffusion may have critical importance.
Furthermore, since the diffusion analysis may be time-critical, or increasing its coverage may be expensive in terms of computing resources and time spent, novel and efficient approaches to find good vertex sets that propagate the information/influence effectively are essential.

Influence maximization is the problem of finding a subset $S \subset V$ of $K$ vertices in a graph $G = (V, E)$ with the vertex set $V$ and edge set $E$ such that $S$ reaches the maximum reachability, i.e., influences the maximum number of vertices, under some diffusion model $M$. Kempe et al.~\cite{kempe2003maximizing} introduced the IM problem, proved it to be NP-hard, and provided a greedy Monte-Carlo approach that has a constant approximation ratio over the optimal solution. This greedy approach is one of the most frequently applied algorithms for IM. The time complexity of the greedy algorithm, estimating the $\sigma$ influence score, running $R$ simulations, and selecting $K$ seed vertices is $\mathcal{O}(KRn\sigma)$ for a graph with $n$ vertices. Although they perform well in terms of seed-set quality, the greedy Monte-Carlo solutions are impractical for real-life networks featuring millions of vertices as a consequence of their expensive simulation costs. Due to this reason, many heuristics and proxy methods have been proposed in the literature~\cite{MixGreedy, narayanam2010shapley, kimura2007extracting, chen2010PMIA,chen2010LDAG, kim2013scalable, cohen2014sketch, goyal2011simpath, jung2012irie,cheng2014imrank,liu2014influence,galhotra2016holistic}.

@@

Simulating a greedy algorithm in parallel is a straightforward workaround to reduce the execution time of IM kernels and make them scalable for large-scale networks. However, for large networks, a greedy approach with a good approximation guarantee does not come cheap even if a large number of processing units/cores are available on networks with billions of vertices and edges. Following similar attempts in the literature, we propose a parallel, sketch-based greedy approach that approximates the Monte-Carlo processes. The proposed approach does not exactly count the number of influenced vertices. Instead, it leverages Count-Distinct sketches. Below is a summary of our contributions:

@@

\item Running concurrent simulations on per-vertex Count-Distinct sketches reduces the number of memory accesses. While traversing an edge, any number of diffusion simulations can be performed between vertices, using only a single (8-bit) value per vertex for each simulation.

\acro processes all samples of a single edge together. The suggested approach, therefore, decreases the pressure on the memory management unit. On the cascade models, \acro employs vector compute units to its near maximum efficiency to regularize memory accesses.

@@

\item We evaluate the performance, memory consumption, and influence score with sketch- and approximation-based state-of-the-art influence maximization algorithms, namely {\sc Skim}~\cite{cohen2014sketch}, {\sc Tim+}~\cite{tim} and {\sc Imm}~\cite{minutoli2019fast}, to accurately position the performance of \acro{} in the IM literature. The experiments show that \acro is $\minspeedup\times$—$\maxspeedup\times$ faster than state-of-the-art sketch-based algorithms while having almost the same quality with the approximation-based algorithms in terms of scores, and using a comparable amount of memory.

@@

The paper is organized as follows: 
In Section~\ref{sec:background}, we present 
the background on IM and introduce the mathematical notation. 
Section~\ref{sec:method} describes the proposed approach in detail.
In Section~\ref{sec:evaluation}, a thorough performance comparison over the traditional algorithms is provided by conducting experiments on various real-world datasets and influence settings. A detailed comparison with the state-of-the-art from the literature is also given.
Section~\ref{sec:relatedwork} presents a comparative overview of the existing work. Finally, Section~\ref{sec:conclusion} discusses future work and concludes the paper.

@@

Let $G = (V,E)$ be a directed graph where the $n$ vertices in $V$ represent the agents, and $m$ edges in $E$ represent the relations between the agents in $V$.
The neighborhood of a vertex $u \in V$ is denoted as $\Gamma_{G}(v) = \{v: (u,v) \in E\}$. 
A graph $G' = (V',E')$ is a sub-graph of $G$ if $V' \subseteq V$ and $E' \subseteq E$. The diffusion probability on the edge $(u, v) \in G$ is noted as $w_{u,v}$, where $w_{u,v}$ can be determined either by the diffusion model or according to the strength of $u$ and $v$'s relationship in the data.

\begin{table}[!ht]
    \caption{Table of notations}
    \label{tab:notation}
    \centering
    \begin{tabular}{|l|p{0.7\linewidth}|}
        \hline
        Variable & Definition  \\
        \hline
        $G = (V,E)$     & Graph $G$ with vertices $V$ and edges $E$ \\
        $\Gamma_G(v)$   & Neighborhood of incoming edges to vertex $v$ in graph $G$\\ %%FIXME
        $w_{u,v}$       & Probability of $u$ directly influencing $v$ \\
        %$SCC(v) $       & Strongly connected component of vertex $v$\\
        $R_{G}(v)$      & Reachability set of vertex $v$ on graph $G$\\
%        $\overline{R_{G}(v)} & Complement of the vertex set $R_{G}(v)$\\
        \hline\hline
        $S$             & Seed set to maximize influence\\
        $K$             & Size of the seed set\\
        $\mathcal{R}$   & Number of Monte-Carlo simulations performed\\
        $\sigma_{G}(S)$ & The influence score of $S$ in $G$, i.e., expected number of vertices reached from $S$ in $G$\\
        % $\sigma_{G}{(S,v)}$          & Marginal influence gain by adding vertex $v$ to seed set $S$\\
        \hline\hline
        $w_{u,v}$             & Sampling probability for the edge $(u,v)$\\
        $P(s,v)_r $     & Random probability generated for selecting edge vertices $s$ to $v$ in simulation $r$\\
        $h(u,v)$        & Hash function for edge $\{u,v\}$\\
        $h_{max}$       & Maximum value hash function $h$ can return\\
        %$X_r$           & Random number/hash generated for simulation $r$  \\
        \hline\hline
        $e$             & Estimated reachability set size\\
        % $[a, \ldots, a]_B$      & Vector of size $B$, contains all $a$\\
        $M_u[j]$        & $j$th sketch register for vertex $u$\\
        $\varsigma $    & Influence gained before last sketch build\\
        $\sigma $       & Influence Score\\
        $\delta$        & Marginal Gain after last sketch build\\
        $Err_l$         & Local estimation error of the sketch\\
        $Err_g$         & Global estimation error of the sketch\\
        $\epsilon_{g}$    & Global estimation error threshold\\
        $\epsilon_{l}$    & Local estimation error threshold\\ 
        $\epsilon_{c}$    & Non-convergenced vertex threshold\\
      %  $x[[i, j]]$   & Slice of vector  $x$ between indices $i$ and $j$, not including $j$ \\ 
        \hline         
    \end{tabular}
\end{table}
\subsection{Influence Maximization}

Influence Maximization aims to find a seed set $S \subseteq V$ among all possible size $K$ subsets of $V$ that maximizes an {\em influence spread function} $\sigma$  when the diffusion process is initiated from $S$. Although we focus on graphs with bidirectional edges, for IM, the edges can be unidirectional depending on the initial construction. That is although the edges $\{u, v\} \in E$ are assumed to be bidirectional, throughout the diffusion process, we use $(u, v)$~($(v, u)$) as the edge that may be used to influence/activate $v$~($u$) given that $u$~($v$) has already been influenced/activated. This being said, $w_{u,v}$ can be set to zero to make the edge $(v, u)$ practically unidirectional.

\begin{figure}[!ht] 
    \centering
  \subfloat[\small{IC}\label{fig:ic}]{%
       \includegraphics[width=0.37\linewidth]{images/ic.pdf}}
  \subfloat[\small{WC}\label{fig:wc}]{%
        \includegraphics[width=0.37\linewidth]{images/wc.pdf}}
    \\% IMAGES HERE
  \caption{\protect\subref{fig:ic} 
The directed graph $G = (V, E)$ for Independent Cascade with independent diffusion probabilities. 
\protect\subref{fig:wc}
The directed graph was obtained from the directed one by setting the diffusion probabilities of incoming edges to $1 / |\Gamma_G(v)|$ for each vertex $v \in V$. 
  }
  %\label{fig_ic_model} 
  \label{fig:xx} 
\end{figure}
In the literature, {\em independent} and {\em weighted cascade}~(IC and WC), and 
{\em linear threshold}~(LT)~\cite{kempe2003maximizing} are three widely recognized diffusion models for IM.

\begin{itemize}[leftmargin=*]
\item An {\bf Independent Cascade} process performs in iterations and activates a vertex $v$  if one of $v$'s (incoming) edges $(u, v)$ is 
used during the diffusion iteration, which happens with the activation probability $w_{u, v}$, given that $u$ has already been influenced in the previous rounds. The activation probabilities are independent (from each other and previous activations) in the {\em independent cascade} model which we concentrate on in this paper, as in Figure~\ref{fig:ic}.
In theory, there can exist parallel and independent $\{u, v\}$ edges in $E$. In practice, they are merged to a single $\{u,v\}$ edge via preprocessing.

\item The {\bf Weighted Cascade}  model is a variant of the independent cascade that uses the structural properties of vertices to set the edge weights as shown in Figure~\ref{fig:wc}.
The method, as described in~\cite{MixGreedy}, sets $w_{u, v} = 1 / d_v$ where $d_v$ is the number 
of incoming edges of $v$~(which in the original graph is equal to $\Gamma _G(v)$).
Therefore, if $v$ has $\ell$ neighbors activated in the last round, its probability of activation in the new round is $1-( 1-1 / d_v)^\ell$.

@@

\item{\bf Linear threshold} generalizes the independent cascade model and activates the $v$ vertex once the cumulative activation coming from the neighbors of $v$ exceeds the $\theta v$ threshold. 
All $\{u, v\}$~(or $(u, v)$ in the directed graph) edges with active $u$ vertices are taken into account in the process. Vertex $v$ is activated when the total activation probability through these edges exceeds $\theta_v$~\cite{kempe2003maximizing}.  
Therefore, the independent cascade model is a special variant of the linear threshold model with $\theta_v = 0$. 
\end{itemize}

As a generalization of these models, {\bf triggering} has been proposed by Kempe~et~al.~\cite{kempe2003maximizing} to generalize the previous models mentioned above. In this model, each neighbor has a probability to influence the vertex $v$. The diffusion process chooses a random subset of vertices called the {\em triggering set} to activate the vertices at each instance.

The complexity analysis stays consistent for many diffusion models, including {\em Independent Cascade}~(IC), {\em Weighted Cascade}~(WC), and {\em Linear Threshold}~(LT)  models. We concentrate on the independent cascade model in this paper, but the proposed methods are also relevant to other models in the literature.

\subsection{Count-Distinct Sketch}\label{sec:sketch}
The {\em distinct element count} problem focuses on finding the number of distinct elements in a stream with non-distinct elements. Computing the reachability sets of a vertex is a similar problem; the cardinality of all vertices visited is calculated while traversing sample subgraphs from the given vertex. Exact cardinality calculations require memory proportional to the cardinality.

The reachability set of a vertex is a distinct union of all connected vertices' reachability sets. Many other Influence Maximization methods exploit this property to some degree.

The methods based on Reverse Reachability and Bottom-Up Traversal utilize this property directly to merge the reachability sets of connected vertices to estimate vertices influence.
MixGreedy method goes one step further; it utilizes that all vertices in a connected component that have the same reachability set in undirected graphs. Therefore, for a sample graph, all reachability sets can be found in a single graph traversal.

For directed graphs, storing reachability sets for all vertices and merging these sets are infeasible for nontrivial graphs. 
If one-hot vectors are used to store reachability sets for constant insertion time, $O(n^2\mathcal{R})$ bits of memory is required and each merge operation has $O(n)$ time complexity. 
if Disjoints sets are used for storing reachability sets; $(n\bar{\sigma}\mathcal{R})$ memory is required to store all reachability sets, and each merge operation has $O(Ackermann'(\sigma))$ complexity.

Count-distinct sketches have nice properties for such problems; Flajolet–Martin algorithm\cite{flajolet1985probabilistic} can estimate the cardinality of distinct elements in reachability sets with a constant number of registers($J$), and most important, union operation on these sets can be done in a constant number of operations.

Flajolet–Martin algorithm only stores how rare elements are in a multiset. The rarity of the elements is often computed by counting the leading zeros in the item's hash value.  The registers are initialized with zero values in the beginning. Items are added by storing the longest leading zero count. Cardinality estimation can be done by taking 2 to the power of the stored value. Multiple registers and hash functions, $M[j]$ and $h_j$, are commonly used to reduce variance.
Adding the item $x$ to sketch $M$, and its result $M'$ is shown in equation~\ref{eq:sketch-add}.

\begin{equation}
    \label{eq:sketch-add}
    M'[j] = max(M[j],clz(h_j(x)) ~ \forall j\in(0,J]
\end{equation} where $clz$ is count leading zeros function and $J$ is the number of sketch registers.

The merge operation for two sketches, $M_u$,$M_v$, can be performed by computing the pairwise maximum of the registers. Formally; 
\begin{equation}
    \label{eq:sketch-merge}
    M'[j] = max(M_u[j],M_v[j]) ~ \forall j\in(0,J]
\end{equation} 
where $M'$ is union of set and $J$ is the number of sketch registers.

Average of the longest running zeros can be used compute the cardinality and the result is divided to the correction factor $\phi \approx 0.77351$ to correct hash collisions.
\begin{equation}
    \label{eq:sketch-estimate}
    e = 2^{\bar{M}}/\phi
\end{equation} 
where $e$ distinct cardinality and $\bar{M}$ is the mean of the registers.

In this work, we utilize a variant of Flajolet–Martin sketch; since multiple Monte-Carlo simulations are performed to calculate the estimated influence, we use one register per simulation and take the average of longest leading zero counts to calculate the average cardinality of reachability sets. Merge operations are performed if and only if there is an edge between the respective simulations.

@@

\kktodo{burada bir giris gerekiyor - basit, 4-5 adimli bir algoritma, bir iki paragraf anlatim ne yapiyoruz, niye yapiyoruz}

\subsection{Hash-based Fused Sampling}
The probabilistic nature of cascade models requires sampling subgraphs $\hat{G}$ from $G = (V, E)$ to simulate the diffusion process. The sampling part of the algorithm can be a dominating factor in processing; sampling may demand multiple passes on the graph, if samples are memoized, multiple times of the graph size needs to be stored.
In this work, we borrow hash-based fused sampling from {\sc InFuseR}\cite{infuser}. By using a hash-based fused sampling, we eliminate the necessity of in-memory creation and storage of the sample subgraphs.
While processing an edge, it is considered for all possible samples.
It is sampled or skipped depending on the outcome of the hash-based random value. The hash function used is given in equation \ref{eq:hash}.
\begin{equation}
    \label{eq:hash}
    h(u,v) = \mbox{{\sc Murmur3}}(u||v)~mod~2^{31}  
\end{equation}
where $||$ is the concatenation operator.

We have tried a few other hash algorithms as well; Ideally, we chose {{\sc Murmur3}} due to its simplicity and good avalanche behavior with maximum bias $0.5 \% $\cite{MurmurHash3Performance}.

For performance considerations, edge hash values were precomputed in our work and only existing pairs were considered. The trade-off between extra memory and computation here maybe not be applicable for different computation architectures and faster/simpler hash functions.

Although the above-mentioned approach generates a unique hash value for each edge, and hence a unique sampling probability, different simulations require different probabilities.

>>>>

\IEEEtitleabstractindextext{%
\begin{abstract}
Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation algorithms are frequently used for IM. In addition to these high-quality, yet expensive approximation algorithms, lightweight, sketch-based approaches, which do not simulate the process exactly, have been proposed in the literature to cope with the scale of today's networks. In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and hash-based fused sampling. To estimate the number of influenced vertices throughout a diffusion, we use per-vertex Flajolet-Martin sketches where each sketch corresponds to a sampled subgraph. To efficiently simulate the diffusions, the reach-set cardinalities of a single vertex are stored in memory in a consecutive fashion. This allows the proposed algorithm to estimate the number of influenced vertices in a single simulation step of different simulations at once. In addition, thanks to their efficiency, and the scalability of our parallel implementation, we can rebuild the sketches whenever it is necessary. For a faster IM kernel, we rebuild the sketches only after observing estimation errors above a given threshold. Our experimental results show that the proposed method yields high-quality seed sets while being up-to $1.7\times$--$27.9\times$ faster on average than a state-of-the-art approximation algorithm. In addition, it is $3.5\times$--$11.1\times$ faster than a sketch-based approach while producing seed sets with $3\%$--$12\%$ better influence scores.
\end{abstract}
}

@@

Efficient information/influence dissemination in a network is an important research area with several applications in various fields, such as viral marketing~\cite{leskovec2007dynamics, trusov2009effects}, social media analysis~\cite{zeng2010social, moreno2004dynamics}, and recommendation systems~\cite{lu2012recommender}.
As the study of these networks is imperative for educational, political, economic, and social purposes, a high-quality seed set to initiate the diffusion may have vital importance.
Furthermore, since the diffusion analysis may be time-critical, or increasing the influence coverage may be too expensive, novel and efficient approaches to find good vertex sets that propagate the information/influence effectively are essential.

Influence maximization is the problem of finding a subset $S \subset V$ of $K$ vertices in a graph $G = (V, E)$ with the vertex set $V$ and edge set $E$ such that $S$ reaches the maximum reachability, i.e., influences the maximum expected number of vertices, under some diffusion model $M$. Kempe et al.~\cite{kempe2003maximizing} introduced the IM problem, proved it to be NP-hard, and provided a greedy Monte-Carlo approach that has a constant approximation ratio over the optimal solution. This greedy approach is one of the most frequently applied algorithms for IM. The time complexity of the greedy algorithm, estimating the $\sigma$ influence score, running $R$ simulations, and selecting $K$ seed vertices is $\mathcal{O}(KRn\sigma)$ for a graph with $n$ vertices. Although they perform well in terms of seed-set quality, the greedy Monte-Carlo solutions are impractical for real-life networks featuring millions of vertices as a consequence of their expensive simulation costs. Due to this reason, many heuristics and proxy methods have been proposed in the literature~\cite{MixGreedy, narayanam2010shapley, kimura2007extracting, chen2010PMIA,chen2010LDAG, kim2013scalable, cohen2014sketch, goyal2011simpath, jung2012irie,cheng2014imrank,liu2014influence,galhotra2016holistic}.

@@

Simulating a greedy algorithm in parallel is a straightforward workaround to reduce the execution time of IM kernels and make them scalable for large-scale networks. However, for large networks, a greedy approach with a good approximation guarantee does not come cheap on networks with billions of vertices and edges even if a large number of processing units/cores are available. Following similar attempts in the literature, we propose a parallel, sketch-based greedy approach that approximates the Monte-Carlo processes. To further increase the performance, the proposed approach does not exactly count the number of influenced vertices. Instead, it leverages Count-Distinct sketches. Below is a summary of our contributions:

@@

\item Running concurrent simulations on per-vertex Count-Distinct sketches reduces the number of memory accesses which is the main bottleneck for many graph kernels in the literature. While traversing an edge, any number of diffusion simulations can be simultaneously performed over the graph, using only a single (8-bit) value per vertex for each simulation.

@@

\item Once it is read from the memory, \acro processes all samples of a single edge together. The suggested approach, therefore, decreases the pressure on the memory subsystem. Furthermore, it employs vector compute units to its near maximum efficiency to regularize memory accesses.

\item We evaluate the performance, memory consumption, and influence score with sketch- and approximation-based state-of-the-art influence maximization algorithms, namely {\sc Skim}~\cite{cohen2014sketch}, {\sc Tim+}~\cite{tim} and {\sc Imm}~\cite{minutoli2019fast}, to accurately position the performance of \acro{} within the IM literature. The experiments show that \acro is $\minspeedup\times$—$\maxspeedup\times$ faster than a state-of-the-art sketch-based algorithm while having almost the same quality with the approximation-based algorithms in terms of influence, and using a comparable amount of memory.

@@

The paper is organized as follows: 
In Section~\ref{sec:background}, we present 
the background on IM and introduce the mathematical notation. 
Section~\ref{sec:method} describes the proposed approach in detail.
In Section~\ref{sec:evaluation}, a thorough performance evaluation is provided by conducting experiments on various real-world datasets and influence settings. A detailed empirical comparison with the state-of-the-art from the literature is also given. Section~\ref{sec:relatedwork} presents a comparative overview of the existing work. Finally, Section~\ref{sec:conclusion} discusses future work and concludes the paper.

@@

Let $G = (V,E)$ be a directed graph where the $n$ vertices in $V$ represent the agents, and $m$ edges in $E$ represent the relations among them. An edge $(u,v) \in E$ is an {\em incoming} edge for $v$ and an {\em outgoing} edge of $u$. The {\em incoming} neighborhood of a vertex $v \in V$ is denoted as $\Gamma_{G}(v) = \{u: (u,v) \in E\}$. A graph $G' = (V',E')$ is a sub-graph of $G$ if $V' \subseteq V$ and $E' \subseteq E$. The diffusion probability on the edge $(u, v) \in G$ is noted as $w_{u,v}$, where $w_{u,v}$ can be determined either by the diffusion model or according to the strength of $u$ and $v$'s relationship in the data.

\begin{table}[!ht]
    \caption{Table of notations}
    \label{tab:notation}
    \centering
    \begin{tabular}{|l|p{0.7\linewidth}|}
        \hline
        Variable & Definition  \\
        \hline
        \kktodo{uzerinden 
        gecelim}\\
        $G = (V,E)$     & Graph $G$ with vertices $V$ and edges $E$ \\
        $\Gamma_G(u)$   & Neighborhood of $u$ in graph $G$, the endpoints of its outgoing edges\\ %%FIXME
        $w_{u,v}$       & Probability of $u$ directly influencing $v$ \\
        %$SCC(v) $       & Strongly connected component of vertex $v$\\
        $R_{G}(v)$      & Reachability set of vertex $v$ on graph $G$\\
%        $\overline{R_{G}(v)} & Complement of the vertex set $R_{G}(v)$\\
        \hline\hline
        $S$             & Seed set to maximize influence\\
        $K$             & Size of the seed set\\
        $\mathcal{R}$   & Number of Monte-Carlo simulations performed\\
        $\sigma_{G}(S)$ & The influence score of $S$ in $G$, i.e., expected number of vertices reached from $S$ in $G$\\
        % $\sigma_{G}{(S,v)}$          & Marginal influence gain by adding vertex $v$ to seed set $S$\\
        \hline\hline
        $w_{u,v}$             & Sampling probability for the edge $(u,v)$\\
        $P(s,v)_r $     & Random probability generated for selecting edge vertices $s$ to $v$ in simulation $r$\\
        $h(u,v)$        & Hash function for edge $\{u,v\}$\\
        $h_{max}$       & Maximum value hash function $h$ can return\\
        %$X_r$           & Random number/hash generated for simulation $r$  \\
        \hline\hline
        $e$             & Estimated reachability set size\\
        % $[a, \ldots, a]_B$      & Vector of size $B$, contains all $a$\\
        $M_u[j]$        & $j$th sketch register for vertex $u$\\
        $\varsigma $    & Influence gained before last sketch build\\
        $\sigma $       & Influence Score\\
        $\delta$        & Marginal Gain after last sketch build\\
        $Err_l$         & Local estimation error of the sketch\\
        $Err_g$         & Global estimation error of the sketch\\
        $\epsilon_{g}$    & Global estimation error threshold\\
        $\epsilon_{l}$    & Local estimation error threshold\\ 
        $\epsilon_{c}$    & Non-convergenced vertex threshold\\
      %  $x[[i, j]]$   & Slice of vector  $x$ between indices $i$ and $j$, not including $j$ \\ 
        \hline         
    \end{tabular}
\end{table}
\subsection{Influence Maximization}

Influence Maximization aims to find a seed set $S \subseteq V$ among all possible size $K$ subsets of $V$ that maximizes an {\em influence spread function} $\sigma$  when the diffusion process is initiated from $S$. %Although we focus on graphs with bidirectional edges, for IM, the edges can be unidirectional depending on the initial construction. That is although the edges $\{u, v\} \in E$ are assumed to be bidirectional, throughout the diffusion process, we use $(u, v)$~($(v, u)$) as the edge that may be used to influence/activate $v$~($u$) given that $u$~($v$) has already been influenced/activated. This being said, $w_{u,v}$ can be set to zero to make the edge $(v, u)$ practically unidirectional. 
In the literature, {\em independent} and {\em weighted cascade}~(IC and WC), and 
{\em linear threshold}~(LT)~\cite{kempe2003maximizing} are three widely recognized diffusion models for IM.

\begin{figure}[!ht] 
    \centering
  \subfloat[\small{IC}\label{fig:ic}]{%
       \includegraphics[width=0.37\linewidth]{images/ic.pdf}}
  \subfloat[\small{WC}\label{fig:wc}]{%
        \includegraphics[width=0.37\linewidth]{images/wc.pdf}}
    \\% IMAGES HERE
  \caption{\protect\subref{fig:ic} 
\kktodo{a parti nonuniform weight olsun} The directed graph $G = (V, E)$ for Independent Cascade with independent diffusion probabilities. 
\protect\subref{fig:wc}
The directed graph for WC is obtained by setting the diffusion probabilities of incoming edges to $1 / |\Gamma_G(v)|$ for each vertex $v \in V$. \kktodo{bu (4,1) kenarinin agirligi 1 degil mi?} 
  }
  %\label{fig_ic_model} 
  \label{fig:xx} 
\end{figure}

\begin{itemize}[leftmargin=*]
\item The {\bf Independent Cascade} model works in rounds and activates a vertex $v$ in the current round if one of $v$'s incoming edges $(u, v)$ is 
used during the diffusion round, which happens with the activation probability $w_{u, v}$, given that $u$ has already been influenced in the previous rounds. The activation probabilities are independent (from each other and previous activations) in the {\em independent cascade} model which we concentrate on this paper. A toy graph with activation probabilities on the edges is shown in Figure~\ref{fig:ic}.
In theory, there can exist parallel and independent $\{u, v\}$ edges in $E$. In practice, they are merged to a single $\{u,v\}$ edge via preprocessing.

\item The {\bf Weighted Cascade}  model is a variant of the independent cascade that uses the structural properties of vertices to set the edge weights as shown in Figure~\ref{fig:wc}.
The method, as described in~\cite{kempe2003maximizing}, sets $w_{u, v} = 1 / d_v$ where $d_v$ is the number 
of incoming edges of $v$~(which in the original graph is equal to $\Gamma _G(v)$).
Therefore, if $v$ has $\ell$ neighbors activated in the last round, its probability of activation in the new round is $1-( 1-1 / d_v)^\ell$. 
\kktodo{bu tanim [7]'dekine uymuyor. Orada undirected graph var } Note that the model is originally defined for undirected graphs in~\cite{MixGreedy}. However, the definition above is consistent since one can consider each~(undirected) edge as two directed edges.

@@

\item{\bf Linear threshold} generalizes the independent cascade model and activates the vertex $v$ once the cumulative activation coming from its neighbors exceeds a given threshold $\theta_v$. 
All the $(u, v)$ edges with active $u$ vertices are taken into account in the process. Vertex $v$ is activated when the total activation probability through these edges exceeds $\theta_v$~\cite{kempe2003maximizing}.  
Therefore, the independent cascade model is a special variant of the linear threshold model with $\theta_v = 0$. 
\end{itemize}

%As a generalization of the above-mentioned models, {\bf triggering} has also been proposed by Kempe~et~al.~\cite{kempe2003maximizing}. In this model, each neighbor has a probability to influence the vertex $v$. The diffusion process chooses a random subset of vertices called the {\em triggering set} to activate the vertices at each instance.
The complexity analysis stays consistent for many diffusion models, including {\em Independent Cascade}~(IC), {\em Weighted Cascade}~(WC), and {\em Linear Threshold}~(LT)  models; the time complexity of the greedy algorithm, estimating the $\sigma$ influence score, running $R$ simulations, and selecting $K$ seed vertices is $\mathcal{O}(KRn\sigma)$ for a graph with $n$ vertices. \kktodo{bu son cumle dogru mu}\ggx{Aklima tersini dogrulayacak bir ornek gelmiyor hocam. Dogru olmali.} We concentrate on the IC model in this paper, but the proposed methods are also relevant to other models in the literature.

\subsection{Count-Distinct Sketches}\label{sec:sketch}
The {\em distinct element count} problem focuses on finding the number of distinct elements in a stream where the elements are coming from a universal set ${\cal U}$. Finding the number of vertices to be influenced of a candidate seed vertex, i.e., the cardinality of the {\em reachability set}, is a similar problem. For each sample subgraph, the number of visited vertices is calculated while traversing the subgraphs starting from the candidate vertex. Note that an exact computation of set cardinality requires memory proportional to the cardinality which is ${\mathcal O}(|{\cal U}|)$.

The reachability set of a vertex is the union of those of all its connected vertices (via outgoing edges). Many IM kernels exploit this property to some degree. The methods based on {\em reverse reachability} \kktodo{cite} and {\em bottom-up traversal} \kktodo{cite} utilize this property directly to merge the reachability sets of connected vertices to estimate the number of vertices influenced. {\em MixGreedy} \kktodo{cite} goes one step further; it utilizes the fact that for an undirected graph, all vertices in a connected component have the same reachability set. Therefore, all the reachability sets within a single sample subgraph can be found via a single graph traversal.

For directed graphs, storing reachability sets for all vertices and merging these sets are infeasible for nontrivial graphs. 
If one-hot vectors are used to store the reachability sets for constant insertion time $\mathcal{O}(n^2\mathcal{R})$ bits of memory is required where each merge operation has $\mathcal{O}(n)$ time complexity. 
If disjoints sets are used for storing reachability sets; $\mathcal{O}(n\bar{\sigma}\mathcal{R})$ \kktodo{Gokhan bunu aciklayalim, bir de $\bar{\sigma}$ nedir}\ggx{estimated influence} memory is required to store all reachability sets, and each merge operation has $\mathcal{O}({\tt Ack}(\sigma))$ complexity where ${\tt Ack}$ is the Ackermann function \kktodo{bu Ackermann'a atif verelim}.

Count-Distinct Sketches can be leveraged to efficiently estimate the cardinality of reachability sets; for instance, the Flajolet–Martin~(FM) sketch~\cite{flajolet1985probabilistic} can do this with a constant number, $J$, of registers. Furthermore, the union of two sketches can be computed in constant time. The FM sketch simply stores that how rare the elements are in a stream. The rarity of the elements is estimated by counting the maximum number of leading zeros in the hash values of the stream elements. Initially, each register is initialized with zero. The items are hashed one by one and the length of the longest all-zero prefix is stored in the register. With a single register, the cardinality estimation can be done by computing the power $2^\ell$ where $\ell$ is the value in the register.

In practice, multiple registers and hash values, $M[j]$ and $h_j$, are commonly used to reduce variance. For a sketch with multiple registers, the impact of adding an item $x \in \cal{U}$ is shown in~\eqref{eq:sketch-add}:
\begin{equation}
    \label{eq:sketch-add}
    M[j] = \max(M[j], clz(h_j(x)),~1 \leq j \leq J
\end{equation} where $clz(y)$ returns the number of leading zeros in $y$ and $J$ is the number of sketch registers. With multiple registers, the average of the register values can be used to estimate the cardinality, and the result is divided to a correction factor $\phi \approx 0.77351$ to fix the error due to hash collisions. That is the estimated cardinality $e$ is computed as
\begin{equation}
    \label{eq:sketch-estimate}
    e = 2^{\bar{M}}/\phi
\end{equation} 
where $\bar{M} = {\tt{avg}}_j\{M[j]\}$ is the mean of the register values.

In this work, we utilize a variant of Flajolet–Martin sketch; since multiple Monte-Carlo simulations are performed to calculate the estimated influence, we use one register per simulation and take the average length of the longest leading zeros. Two given FM sketches $M_u$ and $M_v$ can be merged, i.e., their union $M_{uv}$ can be computed, by taking the pairwise maximums of their registers. Formally; 
\begin{equation}
\label{eq:sketch-merge}
    M_{uv}[j] = \max(M_u[j], M_v[j]),~1 \leq j \leq J.
\end{equation} 
In our implementation, the merge operations are performed if and only if there is an edge between the vertices in respective samples \kktodo{burada simulations yaziyordu, anlamadim, iki simulation arasinda nasil edge olur - o yuzden vertex yazdim}. \ggx{ j'th sampleda vertexler arasında sampled edge olmasını kastetmiştim. Yanlis olmus.}

@@

Most IM algorithms have the same few steps to find the best seed vertex set; sampling, building the influence oracle, verifying the impact of new candidates, then removing the residual reachability set of the latent seed set. Following the idea proposed in~\cite{infuser}, \acro fuses the sampling step with other steps to avoid reading the graph multiple times.

\acro first performs a diffusion process; the process starts with per-vertex sketches that are initialized with the hash value of the corresponding vertex~(i.e., every vertex reaches to itself). Then, for all the (sampled) edges ($u, v$), i.e., the ones that contribute to the diffusion for this simulation, the sketch of the source vertex $u$ is merged into that of the target vertex $v$ until all sketch registers for all the vertices converge. The merge operation utilized in this process is slightly different than the conventional one and retrofitted to mimic the IC diffusion.
